### 2021-08-15 doc/thesis/chapters/system-architecture-and-concept.tex
all /ref to ~/ref (to avoid newline separations)

\noindent
A very high-level overview of the entire system is given by the block diagram in Figure \ref{fig:bd_system_overview_highlevel}.
It shows the front end, which processes the received antenna signal, so that it can be used in the FM radio receiver block.
The radio receiver itself is implemented in multiple different variants.
The produced audio output is sent to a speaker, since the ultimate goal is to listen to a radio stations' audio signal.

@@

In the context of a system level design approach, several steps are necessary to find a solution that is ready to be implemented.
As a very first step, the inner workings of an FM receiver need to be understood.
This mainly regards the signal processing theory, to be able to decode the signal that is received by the antenna.
Chapter \ref{cha:SignalProcessingTheory} describes this part into detail.

@@

\begin{itemize}
  \item \textbf{GNU Radio}\\
      GNU Radio is a free and open-source software development toolkit to implement a software-defined radio.
      The implementation is done by connecting functional blocks in a block design.
      This abstracts the inner workings of each functional block and thus provides an implementation approach on a very high level of abstraction.
      GNU Radio supports interfaces to external RF hardware \cite{SoftwareGnuRadio}.
  \item \textbf{Matlab}\\
      Matlab is a software platform that can work with matrices and compute numerical solutions.
      Matlab also includes various toolboxes, that support developers with complex calculations and tasks on datasets.
      This includes a range of toolboxes for signal processing, such as filter designers or FFT functions, as well as convenient ways to visualize data \cite{SoftwareMatlab}.
  \item \textbf{FPGA: \cplusplus\ High-Level Synthesis}\\
      In order to use an FPGA, a hardware design needs to be developed for it.
      This can be done using high-level synthesis.
      This thesis focuses in \cplusplus and does not implement a SystemC variant, because Xilinx has deprecated SystemC support for synthesis in their latest HLS tools.
      This, and the details of High-level synthesis (HLS) are explained in Chapter \ref{cha:HLS}.

@@

An example of a system concept and its architecture is developed and shown in Figure \ref{fig:bd_system_overview_detailed}.
Here, a common data source is provided by the block called \textit{Front End}.
It records data from an antenna and saves it to a file, in a format that the consequent blocks require at their input.
Therefore, a reproducable input is ensured.
Within the \textit{FM Radio Receiver} block, all the different implementation variants are shown in parallel.
They all work on this common input source and generate their respective output.
This output data is again written to a file.
Hereby it is important to note, that the output of the \textit{Matlab} implementation is chosen to be the 'golden reference model' of correct output.
Its output is used as the verification data later in the process.
This is because Matlab was used to research decoding algorithms and to develop the DSP chain, which is used in the GNU Radio, HLS and VHDL implementations.
In addition to the verification data, the Matlab model also writes common parameters to a file.
These parameters are directly used in the other implementations, to ensure a common setup.
The final block is the \textit{Verification} entity, which compares the output files of all the variants, with respect to the chosen model parameters.\\

>>>>

\noindent
A very high-level overview of the entire system is given by the block diagram in Figure~\ref{fig:bd_system_overview_highlevel}.
It shows the front end, which processes the received antenna signal, so that it can be used in the FM radio receiver block.
The radio receiver itself is implemented in multiple different variants.
The produced audio output is sent to a speaker, since the ultimate goal is to listen to a radio stations' audio signal.

@@

In the context of a system level design approach, several steps are necessary to find a solution that is ready to be implemented.
As a very first step, the inner workings of an FM receiver need to be understood.
This mainly regards the signal processing theory, to be able to decode the signal that is received by the antenna.
Chapter~\ref{cha:SignalProcessingTheory} describes this part into detail.

@@

\begin{itemize}
  \item \textbf{GNU Radio}\\
      GNU Radio is a free and open-source software development toolkit to implement a software-defined radio.
      The implementation is done by connecting functional blocks in a block design.
      This abstracts the inner workings of each functional block and thus provides an implementation approach on a very high level of abstraction.
      GNU Radio supports interfaces to external RF hardware \cite{SoftwareGnuRadio}.
  \item \textbf{Matlab}\\
      Matlab is a software platform that can work with matrices and compute numerical solutions.
      Matlab also includes various toolboxes, that support developers with complex calculations and tasks on datasets.
      This includes a range of toolboxes for signal processing, such as filter designers or FFT functions, as well as convenient ways to visualize data \cite{SoftwareMatlab}.
  \item \textbf{FPGA: \cplusplus\ High-Level Synthesis}\\
      In order to use an FPGA, a hardware design needs to be developed for it.
      This can be done using high-level synthesis.
      This thesis focuses in \cplusplus and does not implement a SystemC variant, because Xilinx has deprecated SystemC support for synthesis in their latest HLS tools.
      This, and the details of High-level synthesis (HLS) are explained in Chapter~\ref{cha:HLS}.

@@

An example of a system concept and its architecture is developed and shown in Figure~\ref{fig:bd_system_overview_detailed}.
Here, a common data source is provided by the block called \textit{Front End}.
It records data from an antenna and saves it to a file, in a format that the consequent blocks require at their input.
Therefore, a reproducable input is ensured.
Within the \textit{FM Radio Receiver} block, all the different implementation variants are shown in parallel.
They all work on this common input source and generate their respective output.
This output data is again written to a file.
Hereby it is important to note, that the output of the \textit{Matlab} implementation is chosen to be the 'golden reference model' of correct output.
Its output is used as the verification data later in the process.
This is because Matlab was used to research decoding algorithms and to develop the DSP chain, which is used in the GNU Radio, HLS and VHDL implementations.
In addition to the verification data, the Matlab model also writes common parameters to a file.
These parameters are directly used in the other implementations, to ensure a common setup.
The final block is the \textit{Verification} entity, which compares the output files of all the variants, with respect to the chosen model parameters.\\

### 2021-08-15 doc/thesis/chapters/system-design-process.tex
all /ref to ~/ref (to avoid newline separations)

Each of these terms represents a separate step in the process diagram shown in Figure \ref{fig:similar_process}.
The steps can be seen as iterative and parallel, as the process runs during the product development cycle.

@@

A well-known process model is the so called V-Model.
It is based on the similarly well-known Waterfall Model, and has its name because of the distinct V-shape of its process diagram, as it is shown in Figure \ref{fig:v_modell}.
One of the main characteristics of the V-Model is, that one phase needs to be completed before the following phase can start.
This classifies the model as a static process model and is one of the main disadvantages at the same time, because it makes the model inflexible to adaptions of the requirements.
The advantages however are that it is easy to install and apply, each phase has its specific deliverables, good results can be achieved for large projects with simple requirements and the early test phases.\\

@@

Furthermore, the V-model diagram in Figure \ref{fig:v_modell} shows the difference between validation and verification, as it is explained in the previous Section \ref{sec:ValidationAndVerification}.

@@

\noindent
HLS is described into more detail in Chapter \ref{cha:HLS}.

@@

Figure \ref{fig:simulink_system_generator} shows this entire design approach in a block diagram.
The abbreviation ISE stands for Integrated Software Environment, and is Xilinx' development environment that includes the FPGA synthesis tool.
ModelSim is a HDL simulator software, that is developed by Mentor Graphics.

@@

Multiple implementation methods are introduced in above sections, which all have different levels of abstraction.
VHDL and Verilog require, or allow, a very detailed description of the RTL code, while SystemC, \cplusplus, or Matlab and Simulink provide a high level of abstraction.
In Figure \ref{fig:fpga_design_levels}, the various implementation methods are compared.
The respective level of abstraction is higher towards the upper end, while the level of detail increases towards the bottom of the figure.\\

@@

%TODO: try to get this figure up into the last page...

>>>>

Each of these terms represents a separate step in the process diagram shown in Figure~\ref{fig:similar_process}.
The steps can be seen as iterative and parallel, as the process runs during the product development cycle.

@@

A well-known process model is the so called V-Model.
It is based on the similarly well-known Waterfall Model, and has its name because of the distinct V-shape of its process diagram, as it is shown in Figure~\ref{fig:v_modell}.
One of the main characteristics of the V-Model is, that one phase needs to be completed before the following phase can start.
This classifies the model as a static process model and is one of the main disadvantages at the same time, because it makes the model inflexible to adaptions of the requirements.
The advantages however are that it is easy to install and apply, each phase has its specific deliverables, good results can be achieved for large projects with simple requirements and the early test phases.\\

@@

Furthermore, the V-model diagram in Figure~\ref{fig:v_modell} shows the difference between validation and verification, as it is explained in the previous Section~\ref{sec:ValidationAndVerification}.

@@

\noindent
HLS is described into more detail in Chapter~\ref{cha:HLS}.

@@

Figure~\ref{fig:simulink_system_generator} shows this entire design approach in a block diagram.
The abbreviation ISE stands for Integrated Software Environment, and is Xilinx' development environment that includes the FPGA synthesis tool.
ModelSim is a HDL simulator software, that is developed by Mentor Graphics.

@@

Multiple implementation methods are introduced in above sections, which all have different levels of abstraction.
VHDL and Verilog require, or allow, a very detailed description of the RTL code, while SystemC, \cplusplus, or Matlab and Simulink provide a high level of abstraction.
In Figure~\ref{fig:fpga_design_levels}, the various implementation methods are compared.
The respective level of abstraction is higher towards the upper end, while the level of detail increases towards the bottom of the figure.\\

@@

%TODO: try to get this figure into the upper page...
